---
title: "Assignment 1"
author: "Kari Roynesdal"
date: "9/8/2020"
output: html_document
---
remotes::install_github("walkerke/tidycensus")

```{r results = "hide", message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)

```


```{r load libraries, message=FALSE}
library(tidycensus)
library(tidyverse)
```


# Variables Used

I wanted to explore the relationship between various population variables such as race, citizenship, and age, with property ownership, property value, household income, and whether an individual owned or rented a property. My unit of analysis is person level. I chose variables from ACS- 1 year census from 2018, and wanted to see this data for the population in Massachusetts at the state level.     

1. Race (categorical variable)
2. Specified owner unit (categorical variable)
3. Citizenship status (categorical variable)
4. Age (continuous) 
5. Household Income (continuous)
6. Property Value (continuous)     


```{r property ownership MA data}

propertyownershipMA_data <- get_pums (variables = c("RAC1P",
                                      "SVAL", 
                                      "FCITP", 
                                      "AGEP",
                                      "HINCP", 
                                      "VALP"),
                        state = "MA", 
                        year = 2018, 
                        survey = "acs1",
                        recode = TRUE) %>%
  filter(VALP > 0) %>%
  select(RAC1P, SVAL_label, FCITP_label, AGEP, HINCP, VALP)


```

Display rows of data frame:

```{r display property ownership data}

propertyownershipMA_data

```

My final data set has 46,398 rows, showing an analysis for the entire state of Massachusetts. 

Saved CSV File: 

```{r write csv file}
write_csv(propertyownershipMA_data, "propertyownershipMA_data.csv")
```








