---
title: "Assignment 5"
author: "Kari Roynesdal"
date: "10/07/2020"
output: html_document
---
remotes::install_github("walkerke/tidycensus")
```{r load libraries, message=FALSE, results='hide'}
library(tidycensus)
library(tidyverse)
library(knitr)
library(jtools)
library(interactions)
library(ggstance)
```

```{r, message=FALSE, results='hide'}
Suffolk_tract_household_data <- read_csv("Suffolk_tracts_households.csv")

```

## Initial Regression Model


I decided to adjust some of my variables from last week, so that I could compare the significance of different household types with percent home ownership.

**Dependent Variable**: 

Percent Owner-Occupied Households 


**Independent Variables**:


The majority racial/ethnic identity

The median age of residents

Total population of the tract

Household Types: 

Female Households with No Husband and No Children Under 18 years old

Female Households with No Husband and Children Under 6 years old

Husband and Wife Households with Children Under 6 years old

Households with Two Male Partners with Children Under 18 years old

Households with Two Female Partners with Children Under 18 years old


```{r}
model1 <- lm(pct_own ~ med_age + tot_pop + majority_race_eth + hh_female_nohusband_nochildrenunder18 +  hh_female_nohusband_childrenunder6yr + hh_husband_wife_fam_childrenunder6yr + hh_unmarriedmale_malepartner_childrenunder18 + hh_unmarriedfemale_femalepartner_childrenunder18,
            data = Suffolk_tract_household_data)
summary(model1)
```

### A Log Transformation 

I think that a percentage increase for median age may be more relevant for my model at the neighborhood level. I am going to log-transform my median age variable to see if it improves the model fit. 

```{r}
model2 <- lm(pct_own ~ log(med_age) + tot_pop + majority_race_eth + hh_female_nohusband_nochildrenunder18 +  hh_female_nohusband_childrenunder6yr + hh_husband_wife_fam_childrenunder6yr + hh_unmarriedmale_malepartner_childrenunder18 + hh_unmarriedfemale_femalepartner_childrenunder18,
            data = Suffolk_tract_household_data)
summary(model2)
```
The R-squared value actually decreased slightly, from 0.7322 to 0.7244. However, the coefficient estimate for median age changed from 0.01520 to 0.5483, which seems to improve the model fit. So every time age doubles for a census tract, the percent of owner-occupied households increases by 0.5483.

### Adding an Interaction Term

I want to see how median age affects two of my variables, Female Households with Children under 6 years and Female Households with No Children under 18: 

```{r}
model3 <- lm(pct_own ~ med_age + tot_pop + majority_race_eth + hh_female_nohusband_nochildrenunder18 +  hh_female_nohusband_childrenunder6yr + hh_husband_wife_fam_childrenunder6yr + hh_unmarriedmale_malepartner_childrenunder18 + hh_unmarriedfemale_femalepartner_childrenunder18 + hh_female_nohusband_nochildrenunder18:med_age + hh_female_nohusband_childrenunder6yr:med_age,
            data = Suffolk_tract_household_data)
summary(model3)
```

The negative and statistically significant coefficient for the interaction between Female Households with Children Under 6 years old and Median Age tells me that median age has less of an effect on this variable. The positive and statistically significant coefficient for the interaction between Female Households with No Children under 18 years old tells me that median age has an effect on this variable. 

I will test another model showing all of these variations for my households variables, using an interaction term of median age: 

```{r}
model4 <- lm(pct_own ~ log(med_age) + tot_pop + majority_race_eth + hh_female_nohusband_nochildrenunder18 +  hh_female_nohusband_childrenunder6yr + hh_husband_wife_fam_childrenunder6yr + hh_unmarriedmale_malepartner_childrenunder18 + hh_unmarriedfemale_femalepartner_childrenunder18 + hh_female_nohusband_nochildrenunder18:med_age + hh_female_nohusband_childrenunder6yr:med_age + hh_husband_wife_fam_childrenunder6yr:med_age + hh_unmarriedmale_malepartner_childrenunder18:med_age + hh_unmarriedfemale_femalepartner_childrenunder18:med_age ,
            data = Suffolk_tract_household_data)
summary(model4)
```
Well that doesn't look like it did much, and now I have very little statistical significance for my dependent variables! 


### Graphing the Results 


I will create a graph to compare the R-squared values for all of my models: 

```{r}
ModelFit <- tibble(model = c(1, 2, 3, 4),
                   R_square = c(summary(model1)$adj.r.squared,
                                summary(model2)$adj.r.squared,
                                summary(model3)$adj.r.squared,
                                summary(model4)$adj.r.squared))

ggplot(ModelFit, aes(x = model, y = R_square)) +
  geom_line() +
  scale_x_continuous(name = "",
                   breaks = breaks <- seq(1, 4, by = 1),
                   labels = paste("Model", breaks)) +
  scale_y_continuous(name = "Adjusted R-squared value") +
  theme_bw()
```

It looks like Model 3 is my best fitting model, but is pretty close to Model 1. 



```{r}
coeff_names <- c("Total Population" = "tot_pop",
                 "Median Age" = "med_age",
                 "Majority Race/Ethnicity Hispanic" = "majority_race_ethhispanic",
                 "Majority Race/Ethnicity Asian" = "majority_race_ethnh_asian",
                 "Majority Race/Ethnicity Black" = "majority_race_ethnh_black",
                 "No Majority Race/Ethnicity" = "majority_race_ethNone", 
                 "Female Household No Children < 18 years" = "hh_female_nohusband_nochildrenunder18",
                 "Female Household Children < 6 years" = "hh_female_nohusband_childrenunder6yr",
                 "Husband Wife Household Children < 6 years" = "hh_husband_wife_fam_childrenunder6yr",
                 "Male Partners Household Children < 18 years" = "hh_unmarriedmale_malepartner_childrenunder18",
                 "Female Partners Household Children < 18 years" = "hh_unmarriedfemale_femalepartner_childrenunder18",
                 "Interaction:\nFemale Household No Children < 18 years and Median Age" = "med_age:hh_female_nohusband_nochildrenunder18",
                 "Interaction:\nFemale Household Children < 6 years and Median Age" = "med_age:hh_female_nohusband_childrenunder6yr")

export_summs(model1, model3, 
             error_format = "(p = {p.value})",
             error_pos = "same",
             model.names = c("Initial model", "Preferred model"),
             coefs = coeff_names)
```

### Visualizing Regression Results

I think Model 1 is actually my preferred model, looking at the statistical significance for my variables. I am going to graph Model 1: 

```{r}
coeff_desc <- c("Total Population" = "tot_pop",
                 "Median Age" = "med_age",
                 "Majority Race/Ethnicity Hispanic" = "majority_race_ethhispanic",
                 "Majority Race/Ethnicity Asian" = "majority_race_ethnh_asian",
                 "Majority Race/Ethnicity Black" = "majority_race_ethnh_black",
                 "No Majority Race/Ethnicity" = "majority_race_ethNone", 
                 "Female Household No Children < 18 years" = "hh_female_nohusband_nochildrenunder18",
                 "Female Household Children < 6 years" = "hh_female_nohusband_childrenunder6yr",
                 "Husband Wife Household Children < 6 years" = "hh_husband_wife_fam_childrenunder6yr",
                 "Male Partners Household Children < 18 years" = "hh_unmarriedmale_malepartner_childrenunder18",
                 "Female Partners Household Children < 18 years" = "hh_unmarriedfemale_femalepartner_childrenunder18")


plot_coefs(model1, 
           coefs = coeff_desc,
           plot.distributions = TRUE,
           rescale.distributions = TRUE) +
  scale_x_continuous(name = "Effect of Household Type on Percent Home Ownership")
```

